CRUSE to Safe Cycling in Ireland

### 2023-09-12 cruse-paper.qmd
Add placeholders

---
title: "CRUSE to Safe Cycling in Ireland"
subtitle: An Open Source Methodology to Support Active Travel
bibliography: bibliography.bib
csl: apa.csl
# Uncomment for draft versions:
format: 
  html:
    output-file: index.html
  pdf:
    output-file: cruse-paper.pdf 
  gfm:
    output-file: README.md
number-sections: true

# Uncomment the following line to produce the cover page
# author:
#   - name: Dr. Robin Lovelace
#     email: r.lovelace@leeds.ac.uk
#     corresponding: true
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Dr. Joey Talbot
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Eugeni Vidal Tortosa
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Hussein Mahfouz
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Dan Brennan
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Dr. Suzanne Meade
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Elaine Brick
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Peter Wright
#     affiliations:
#       - name: AECOM
#         address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK
# abstract: |
#   Under the EU Road infrastructure safety management (RISM) directive, the National Road Safety Strategy (RSS), and the Climate Action Plan Transport Infrastructure Ireland (TII) has a remit for road safety and decarbonizing a predominantly road-based network in Ireland.

@@

[@lovelace2017]

• Title page/main information (title, full names and institutional addresses for all
authors, indication of the corresponding author)
• Abstract
• Introduction

@@

>>>>

---
title: "CRUSE to Safe Cycling in Ireland"
subtitle: An Open Source Methodology to Support Active Travel
bibliography: bibliography.bib
csl: apa.csl

# Uncomment for draft versions:
format: 
  docx: default
  html:
    output-file: index.html
  pdf:
    output-file: cruse-paper.pdf 
  gfm:
    output-file: README.md
number-sections: true

# # Uncomment the following line to produce the cover page
# author:
#   - name: Dr. Robin Lovelace
#     email: r.lovelace@leeds.ac.uk
#     corresponding: true
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Dr. Joey Talbot
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Eugeni Vidal Tortosa
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Hussein Mahfouz
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Dan Brennan
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Dr. Suzanne Meade
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Elaine Brick
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Peter Wright
#     affiliations:
#       - name: AECOM
#         address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK
# abstract: |
#   Under the EU Road infrastructure safety management (RISM) directive, the National Road Safety Strategy (RSS), and the Climate Action Plan Transport Infrastructure Ireland (TII) has a remit for road safety and decarbonizing a predominantly road-based network in Ireland.

@@

Transport is a major contributor to climate change, premature death due to road traffic collisions and air pollution, yet has the potential to be a major contributor to decarbonization and improved public health.
Transport is responsible for 23% of global emissions, 70% of which is from road transport, nearly half of which is from passenger cars [@jaramillo2022].
The transport system encourages, enables and in some cases enforces unsustainable lifestyles, including over-consumption of goods due to excessive mobile storage space and dependency on services that are only accessible by car due to land use plans that have built up around roads [@gray2001; @shergold2012; @motte-baumvol2010].

Recognizing the impacts of poorly designed and performing transport systems on their citizens, governments in many countries have set targets and taken actions and set targets.
In the context of climate, road safety and physicial inactivity crises, policies to improve transport systems can be classified according to the 'Avoid-Shift-Improve' (ASI) framework [@jaramillo2022].
The framework highlights the importance of demand reduction (*avoid*ing unnecessary trips), in addition to mode *shift* and *improvement* of existing energy converters, in that order.

Uptake of cycling, the main topic of this paper, should be seen in this broader context of transport decarbonisation.
Although cycling uptake appears on the surface to only relate to the 'shift' part of the ASI framework, closer consideration of the knock-on impacts of cycling uptake shows that it can also help avoid unnecessary trips \[citation needed, any ideas Hussein?\]
.
The rapid uptake of highly efficient e-bikes can also be seen as an improvement on most electric vehicles, which are too heavy and expensive to be a sustainable alternative and could in fact delay the transition away from car dependency \[citation needed\].

At the European Union level ...

In Ireland...

\[Gary, Suzanne, others input here please.\]

@@

Commute and school travel

Other trip purposes

\[Joey please add your bit here.\]

Bit on Jittering [@lovelace2022]
